Modelling the emergence of whisker barrels

### 2020-06-29 paper/briefpaper/brief.tex
Bringing in changes from Stuarts google doc, and preparing to shift as much paramsearch as possible into figure caption

% Colours for revisions:
% Custom colours
\definecolor{revgreen}{rgb}{0.220, 0.463, 0.114}
\definecolor{revblue}{rgb}{0.235, 0.471, 0.847}
\definecolor{revred}{rgb}{1.0, 0.0, 0.0}
\definecolor{revgrey}{rgb}{0.6, 0.6, 0.63}
\definecolor{revcomp}{rgb}{0.0, 0.0, 0.0}
\definecolor{revpurp}{rgb}{0.8, 0.1, 0.8}

% Blue changes are mostly major points 2 (competition mechanism) and 3 (parameter/sensitivity)
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
% Red for Stuart's updates
\newcommand{\stu}[1]{\textcolor{red}{#1}}
% Minor point 4:
\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
% Minor point 6:
\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}

@@

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}, \cmnt{Fig.\,5b}). \textbf{C} Simulation results for parameters
      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \stu{the}
      simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
      line); black squares \cmnt{provide a} measure \cmnt{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta(t)$ (the \cmnt{pattern quality metric---see
        Materials \& Methods; units mm$^3$});
      \cmnt{grey hexagons \stu{provide a measure of how selectively each
          cortical site is innervated};
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      \textbf{E} \stu{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \cmnt{The colour map
        indicates the values of $\mu(\mb{x})$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

First we verified that all results established by \cite{karbowski_model_2004}
for a 1D axis could be reproduced using our extension to a 2D cortical
sheet. Using an elliptical domain, $S$, with $M=3$ offset guidance gradients
aligned to the longer axis, $N=5$ thalamocortical projections gave rise to
five distinct cortical fields at locations that preserved the topographic
ordering defined by the original $\gamma$ values. However, we found that
specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling
fields. This is because localization of axon densities occurs only when
projections are influenced by interactions with two or more signalling
gradients that encourage migration in opposing directions. Note that these
dynamics are quite unlike classic chemospecificity models
\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
i.e., conditions in the target tissue that instruct pre-identified afferents
to stop growing. As the number of distinct guidance fields is unlikely to
approach the number of barrels, further extensions were required.

The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
thalamocortical projections for a limited availability of cortical
connections. To introduce competition also in terms of axon branching,
\cmnt{whilst ensuring that $a_i$ is conserved over time,} we
redefined
%
\begin{equation} \label{eq:comp}
  \color{blue}
  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
  \color{black}
\end{equation}
%
\stu{This term contributes to the \emph{flux of axonal branching} as an
  additional source of diffusion, which reduces the branching density for a
  given projection where the branches of other projections are dense.}
\stu{The strength of this additional competition is parameterized by
  $\epsilon$}. Note that this operation is local to individual afferent
projections.

The only differences between thalamic projections are their strengths of
interaction with the guidance fields, $\gamma$, hence any reliable differences
between emergent cortical fields must be due to differences in these values
only. We speculate that the contribution of a given ephrin field, to the
velocity at which a projection migrates across the cortical subplate, is
determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.

\stu{As such, we consider linear gradients aligned to anterior-posterior and
  medial-lateral axes through the VPM, and as $\gamma$-values use the
  coordinates of the barreloid centers in these axes as they appear in
  \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary
  thalamic and cortical molecular guidance gradients of this form have been
  well established experimentally}
\citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g.,
  cells originating in the medial part of ventrobasal thalamus (comprising VPM
  and VPL) express high levels of EphA receptors and project to the lateral
  part of S1, which expresses low levels of ephrin-A5, and cells originating
  in the lateral part of ventrobasal thalamus express low levels of EphA
  receptors and project to the medial part of S1, which expresses high levels
  of ephrin-A5 (see}
\citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}). \stu{We
  concentrate on the representation of the 41 macrovibrissae because their
  thalamic and cortical correlates are easily identifiable and consistently
  positioned, we exclude the microvibrissae to limit the complexity of the
  simulations, and we exclude the five rhinal whiskers as their representation
  in S1 is isolated from the main barrel field.}

A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using original data kindly supplied by the authors)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and medial-lateral axes.

\mpfour{The signalling gradients are assumed not to interact with the barrel
  field boundary and to continue well beyond its extent (see
  \citealp{miller_epha7-ephrin-a5_2006}, their Fig.\,11) in order to
  participate in other developmental processes, such as inter-area positional
  specification \citep{sur_patterning_2005,oleary_area_2007}.} % Just oleary_area_2007 would be ok
%
From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
  $c(\mb{x},0)=0$}, a clear Voronoi-like tesselation
emerged (Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta
metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
pattern emerged within \cmnt{$\approx$~20000} iterations
(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
between real and simulated barrels revealed a strong correspondence (see
Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
  \emph{the connection density of the TC type $i$ which makes the densest
    connections} to \emph{the sum of all connection densities} gives a measure
  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
  across the entire barrel domain and named it $\omega(t)$ (see
  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
  20000 simulation steps and converges on a stable
  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
  luminance map, giving a more graded impression of the structure which
  develops in the system.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\cmnt{\textbf{A} Exploring the parameter space of the model given
        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
        barrel area difference and $\omega$ (bottom row) is a measure of
        selectivity of the connection densities for the 41
        thalamocortical axon types. For each row, lighter (orange/yellow)
        colours imply a better model. The parameter space explored is three
        dimensional with the parameter ratio $\alpha/\beta$ varied along the
        columns of graphs, the competition parameter $\epsilon$ varied on the
        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
        asterisk (*) marks the parameter set displayed in
        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\cmnt{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
  iterations plotted as colour maps for $216$ parameter sets. Each row of maps
  shows the results for a separate metric. Each column of maps shows
  results for a different value of the ratio $\alpha / \beta$. The diffusion
  parameter, $D$ varies with the $x$ axes and the strength of competition,
  $\epsilon$ varies with the $y$ axes. The colour maps are arranged so that
  `yellow is good'. Lower values of $\delta$ indicate good Voroni patterns and
  are yellow. Lower values of the difference between real and simulated barrel
  area indicates good pattern matching. Higher values of $\omega$ indicate
  greater selectivity and are coloured yellow. White squares indicate that the
  simulation was not stable for that parameter set.}

\cmnt{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the
  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
  trade-off between good selection and good Voroni pattern
  conformation. The
  parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
  of varying $D$ in either direction from its optimum. ii)~is a less
  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig3.png}
    \caption{
      \cmnt{\textbf{A} Sensitivity of the system pattern
        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
        (bottom row), evaluated at $t=30000$ steps,
        to three classes of perturbation. $y$-axes and colour scales have identical ranges
        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
        with the parameter exploration.
        %
        \emph{Column 1:} Random noise drawn from a uniform distribution,
        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
        to the values for the interaction parameters, $\gamma_{i,j}$ used in
        Fig.\,\ref{fig:main}.
        %
        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
        was smoothed by convolution with a symmetric 2D
        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
        thus show the effect of changing the magnitude and the length scale of
        noise in the guidance fields.
        %
        \emph{Column 3:} Colour maps show the result of setting the rotational angle
        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
        parameters of $\rho_2(\mb{x})$
        unchanged from those used in the simulation in
        Fig.\,\ref{fig:main}, for which $\phi_{\rho_1}=6$\textdegree~and
        $\phi_{\rho_2}=84$\textdegree.
        %
        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
        the patterns. i) The effect of significant interaction parameter noise
        causes barrels to become swapped from the positions they would
        otherwise take. ii) High magnitude, short lengthscale noise in
        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
        which reduces the strength of the axons' interaction with the guidance
        field in the left-right sense. Barrel rows B, C and D become `crushed'
        down the center line, and edge barrels dominate. iv) The effect of
        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
        the pattern in the anticlockwise sense.
      }
    }
    \label{fig:sens}
  \end{fullwidth}
\end{figure}

\cmnt{We performed a sensitivity analysis to determine if the quality of the
  pattern was affected by perturbations to any of the following inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
%
  In the following analysis, we always evaluated the three pattern metrics,
  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
  $\omega$ (which measures selectivity of connections, $c_i$) after running
  the model for 30000 simulation steps.}

\cmnt{First, we conducted 10 runs of the simulation shown in
  Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
  three metrics in the default model. The
  results (where deviations occur as a result of the different random initial
  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
  $\omega=0.2165\,(0.0001)$. None of the three metrics were strongly affected
  by the random changes to $a_i$ on each run. The results give a baseline from
  which to evaluate whether changes in the other inputs have a statistically
  significant effect.}

>>>>

% Colours for revisions:
% Custom colours
\definecolor{revgreen}{rgb}{0.220, 0.463, 0.114}
\definecolor{revblue}{rgb}{0.235, 0.471, 0.847}
\definecolor{revred}{rgb}{0.8353, 0.1255, 0.1255}
\definecolor{revorange}{rgb}{0.8235, 0.4275, 0.0}
\definecolor{revgrey}{rgb}{0.6, 0.6, 0.63}
\definecolor{revcomp}{rgb}{0.0, 0.0, 0.0}
\definecolor{revpurp}{rgb}{0.8, 0.1, 0.8}
\definecolor{revpurp2}{rgb}{0.3882, 0.0745, 0.7098}
\definecolor{revpink}{rgb}{0.7098, 0.0745, 0.6431}
\definecolor{revbblack}{rgb}{0, 0, 0.3}
\definecolor{revgblack}{rgb}{0, 0.2, 0}

% Blue changes are major point 2 (competition mechanism) and other, general
% uncategorised changes.
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
% Red for Stuart's first updates
\newcommand{\stu}[1]{\textcolor{red}{#1}}
% Minor point 4:
\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
% Minor point 6:
\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
% Justify competition:
\newcommand{\justifycomp}[1]{\textcolor{revpurp2}{#1}}
% Justify gamma/interaction parameter choice
\newcommand{\justifygammas}[1]{\textcolor{revpink}{#1}}
% Describe the metrics
\newcommand{\metrics}[1]{\textcolor{revred}{#1}}
% Parameter search. Major point 3
\newcommand{\paramsearch}[1]{\textcolor{revbblack}{#1}}
% Sensitivity analysis. Major point 3
\newcommand{\sensitivity}[1]{\textcolor{revgblack}{#1}}

@@

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}, \cmnt{Fig.\,5b}). \textbf{C} Simulation results for parameters
      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \cmnt{the}
      simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
      line); black squares \cmnt{provide a} measure \cmnt{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta(t)$ (the \cmnt{pattern quality metric---see
        Materials \& Methods; units mm$^3$});
      \cmnt{grey hexagons \cmnt{provide a measure of how selectively each
          cortical site is innervated};
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      \textbf{E} \cmnt{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \cmnt{The colour map
        indicates the values of $\mu(\mb{x})$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

First we verified that all results established by \cite{karbowski_model_2004}
for a 1D axis could be reproduced using our extension to a 2D cortical
sheet. Using an elliptical domain, $S$, with $M=3$ offset guidance gradients
aligned to the longer axis, $N=5$ thalamocortical projections gave rise to
five distinct cortical fields at locations that preserved the topographic
ordering defined by the original $\gamma$ values. However, we found that
specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling
fields. This is because localization of axon densities occurs only when
projections are influenced by interactions with two or more signalling
gradients that encourage migration in opposing directions.
%
%Note that these dynamics are quite unlike classic chemospecificity models
%\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
%i.e., conditions in the target tissue that instruct pre-identified afferents
%to stop growing.
%
As the number of distinct guidance fields is unlikely to
approach the number of barrels, \cmnt{modifications to the model} were required.

% Justify competition term. This pulled from Stuart's Google doc
\justifycomp{We reasoned that an arbitrary number of distinct field locations
  may be specified using a minimum of two guidance fields, if the tendency for
  the density of each projection to concentrate is appropriately balanced
  against the tendency of projections that interact more strongly with the
  guidance fields to migrate further in the directions specified by their
  gradients. Accordingly, projections that interact most strongly with a given
  guidance gradient would come to occupy cortical locations at which that
  field has extreme values, leaving adjacent locations available to be
  occupied by projections with the next strongest interactions, and so
  forth. This would in principle allow the \emph{relative} locations of the
  fields to be specified by the relative values of the interaction parameters,
  $\gamma$, and hence for a topological map in the cortex to be specified by a
  spatial ordering of the $\gamma$ values at the level of the thalamus.}

\justifycomp{Such dynamics are quite unlike those described by classic
  chemospecificity models} \citep{sperry_chemoaffinity_1963},
\justifycomp{which essentially assume center-points by specifying conditions
  in the target tissue that instruct pre-identified afferents to stop
  growing. Consider, for example, that when simulated in isolation from
  one-another, all projections in the model described would simply migrate to
  the extrema of the cortical guidance fields.}

\justifycomp{Testing this reasoning required increasing the strength of the
  competition between simulated thalamocortical projections for cortical
  territory, by increasing the tendency for the density of each projection to
  concentrate. The major modification required was thus to introduce into the
  model an additional source of competition between thalamic projections.}
%
The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
thalamocortical projections for a limited availability of cortical
connections. To introduce competition also in terms of axon branching,
\cmnt{whilst ensuring that $a_i$ is conserved over time,} we
redefined
%
\begin{equation} \label{eq:comp}
  \color{blue}
  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
  \color{black}
\end{equation}
%
\justifycomp{This term contributes to the \emph{flux} of axonal branching as
  an additional source of diffusion, which reduces the branching density for a
  given projection where the branches of other projections are dense. The
  strength of this additional competition is parameterized by
  $\epsilon$}. Note that this operation is local to individual afferent
projections.

\justifygammas{In addition, the model outlined requires that molecular
  guidance gradients in the cortex are complemented by graded values of the
  interaction strengths, $\gamma$, at the level of the thalamus. While the
  precise mechanisms by thalamic and cortical gradients interact during
  development have not been fully characterised, the presence of complementary
  thalamic and cortical molecular guidance gradients has been well established
  experimentally. In particular, the EphA4 receptor and its ligand ephrin-A5
  are distributed in complementary gradients in the somatosensory thalamus and
  cortex} (\citealp{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}).
\justifygammas{Cells originating in the medial part of ventrobasal thalamus
  (comprising VPM and VPL) express high levels of EphA receptors and project
  to the lateral part of S1, which expresses low levels of ephrin-A5, and
  cells originating in the lateral part of ventrobasal thalamus express low
  levels of EphA receptors and project to the medial part of S1, which
  expresses high levels of ephrin-A5} (see
\citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).

\justifygammas{For simplicity, both the thalamic interaction gradients and the
  cortical guidance fields in the model were chosen in the first instance to
  be linear and orthogonal. Hence a given pair of gamma values corresponds to
  the coordinate of a barreloid center in the VPM (in a reference plane
  defined by the anterior-posterior and medial-lateral axes. Coordinates were
  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \justifygammas{and
  scaled (arbitrarily) such that $\gamma\in\pm 2$.}

A cortical boundary enclosing \cmnt{the} 41 \cmnt{macrovibrissae} barrels was
traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}
\cmnt{(using original data kindly supplied by the authors)}, and
Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
resulting domain, using $M=2$ linear signalling gradients aligned with the
anterior-posterior and medial-lateral axes.  \mpfour{These gradients may
  extend well beyond the boundary of the barrel field (e.g. as ephrin-5 does;
  see \citealp{miller_epha7-ephrin-a5_2006}, their Fig.\,11) but they are
  shown within the barrel fields boundary in Fig.\,\ref{fig:main}A to aid the
  visualization. Simulations were iterated through 30000 iterations of
  Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta t=0.0001$)}

From initial conditions sample from a uniform random distribution for
$a(\mb{x},0)\in(0.2,0.4)$ and $c(\mb{x},0)=0$, and across a wide range of
parameter values,
%From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
%  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
%  $c(\mb{x},0)=0$},
a clear Voronoi-like tesselation \cmnt{of topographically organised
  projections} emerged, \cmnt{confirming that barrel maps can self-organise in
  the absence of pre-specified center points. Parameters for the example
  simulation shown in Fig.\,\ref{fig:main}C (see Movie S1) were obtained by
  conducting a full parameter sweep and choosing a combination of values that
  scored well against three measures.}

% New section to describe and introduce the metrics
\metrics{First, we used an algorithm introduced by Honda to measure the
  discrepancy of each barrel shape from a Dirichletform shape. Low
  overall values of the \emph{Honda-}$\delta$ metric obtained from simulated
  barrels indicate a close correspondence of the simulated barrel field with a
  Voronoi tesselation, and thus with a biological barrel field. For the
  % Run ./build/sim/honda boundaries/rat_barrels/wb_110405_Dirichlet.svg to
  % get this result:
  tesselation overlaid on the real barrel field in Fig.\,1A, $\delta=0.025$,
  and a reduction in the Honda-$\delta$ metric in simulation
  confirmed that an equivalent `good' Voronoi pattern} (see
    \citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000
  iterations (Fig.\,\ref{fig:main}D, red squares). Second, we devised a
  \emph{pattern difference} measure that is sensitive to deviations in the
  component shapes and overall topographic registration of two tesselations,
  $\eta$, and we used this measure to compare the simulated barrel fields to
  the real barrel field from which the simulation boundary shape was obtained
  (see Materials and Methods for details). A similar reduction in pattern
  difference over simulation time confirmed that the shapes and arrangement of
  emergent connection fields matched those of the real barrel field by around
  20000 iterations. Third, we measured the \emph{connection selectivity} at
  each location on the cortical sheet, as the connection density of the most
  dense projection divided by the sum over all projection densities, $\omega$.}

\metrics{Against these three metrics we are also able to characterise the
  robustness of self-organization to the model parameters, and to investigate
  the sensitivity of the model to variation in its inputs.}

% Most of this commented out region is covered in the para above.
%
%A reduction in the Honda delta
%metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
%pattern emerged within \cmnt{$\approx$~20000} iterations
%(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
%between real and simulated barrels revealed a strong correspondence (see
%Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
%  \emph{the connection density of the TC type $i$ which makes the densest
%    connections} to \emph{the sum of all connection densities} gives a measure
%  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
%  across the entire barrel domain and named it $\omega(t)$ (see
%  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
%  20000 simulation steps and converges on a stable
%  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
%  luminance map, giving a more graded impression of the structure which
%  develops in the system.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the model given
        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
        barrel area difference and $\omega$ (bottom row) is a measure of
        selectivity of the connection densities for the 41
        thalamocortical axon types. For each row, lighter (orange/yellow)
        colours imply a better model. The parameter space explored is three
        dimensional with the parameter ratio $\alpha/\beta$ varied along the
        columns of graphs, the competition parameter $\epsilon$ varied on the
        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
        asterisk (*) marks the parameter set displayed in
        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\paramsearch{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
  iterations plotted as colour maps for 216 parameter sets. Each row of maps
  shows the results for a separate metric. Each column of maps shows results
  for a different value of the ratio $\alpha / \beta$. The diffusion
  parameter, $D$, varies with the $x$ axes and the strength of competition,
  $\epsilon$ varies with the $y$ axes. The colour maps are chosen so that
  yellow indicates desirable properties for the simulations: Lower values of
  $\delta$ indicate good Voroni patterns and are yellow; Lower values of
  $\eta$ indicate good pattern matching; Higher values of $\omega$ indicate
  greater selectivity and are also coloured yellow. White squares indicate
  that the simulation was not stable for that parameter set.}

\paramsearch{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the
  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
  trade-off between good selection and good Voroni pattern
  conformation. The
  parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
  of varying $D$ in either direction from its optimum. ii)~is a less
  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}
